---
title: "DavidLudeke_A2"
author: "David Ludeke"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(tidyverse)
library(sf)
library(tigris) 
library(censusapi)
library(mapview)
library(leaflet)
```

```{r}
# Using developer version of tigris so I can grab 2020 block shapefiles correctly. 
library(devtools) 
install_github('walkerke/tigris')
library(tigris)
```

```{r}
# Requested my own Census API key from the Census. 
Sys.setenv(CENSUS_KEY="46c0518594422fc024772060dae78c2c2ffcc89d")
```

```{r}
# Grab 2020 Census data
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables")
sfc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:075", # State:06 = California, County:075 = San Francisco (my hometown!)
    vars = "P1_001N") %>% # Variable code for TOTAL population
  transmute(
    GEOID20 = paste0(state,county,tract,block),
    pop_2020 = P1_001N # Variable code for TOTAL population
  )

# Prepare 2020 census blocks for my chosen county in California.
blocks_2020 <- blocks("CA", "San Francisco", year = 2020, progress_bar = F)
```

```{r}
# Grab 2010 Census data
dec_vars_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl/", 
    type = "variables"
  )
sfc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:075",
    vars = "P001001") %>% # Variable code for TOTAL population in the 2010 census.
  transmute(
    GEOID10 = paste0(state,county,tract,block),
    pop_2010 = P001001
  )

# Prepare 2010 census blocks for my chosen county in California. 
blocks_2010 <- blocks("CA", "San Francisco", year = 2010, progress_bar = F)
```

Create a map of the blocks for San Francisco County based on the geometries from the 2010 Census blocks. 

```{r}
# Download the Census Designated Place shapefile for San Franciscio
sfc_boundary <- places("CA", cb = T, progress_bar = F) %>% 
  filter(NAME == "San Francisco")

# Define set of SF County blocks from 2010 based on which have centroids within the boundary.
sfc_blocks_2010 <- blocks_2010 %>% 
  st_centroid() %>% 
  .[sfc_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(blocks_2010 %>% select(GEOID10)) %>% 
  st_as_sf() %>% 
  select(GEOID10) %>% 
  left_join(sfc_pop_2010) # Add in population data for each block in 2010

# Create map of boundary for San Francisco blocks 
mapview(sfc_boundary) + mapview(sfc_blocks_2010, alpha.regions = 0, color = "orange", lwd = 2)

```

```{r}
# Get all 2020 census blocks that intersect with the 2010 blocks.
sfc_blocks_2020 <- blocks_2020 %>%
  .[sfc_boundary, ] %>% 
  select(GEOID20) %>% 
  left_join(sfc_pop_2020)
```

```{r}
# Spatial subsetting method shown in Chapter 2.4.
sfc_blocks_2020_intersect <- sfc_blocks_2020 %>% 
  st_transform(26910) %>% 
  mutate(original_area = st_area(.)) %>% 
  st_intersection( # "Cookie-cutter" procedure
    sfc_blocks_2010 %>%
      select(GEOID10) %>% 
      st_transform(26910)
  ) %>% 
  mutate(
    leftover_area = st_area(.),
    perc_area = as.numeric(leftover_area / original_area),
    pop_2020 = pop_2020 * perc_area
  )
```

```{r}
# Group and summzarize to get the 2020 population counts within the GEOID10 boundaries. 
sfc_blocks_2020_reshaped <- sfc_blocks_2020_intersect %>%
  st_drop_geometry() %>% 
  group_by(GEOID10) %>% 
  summarize(
    pop_2020 = sum(pop_2020, na.rm=T) %>% round()
  )
```

```{r}
# Compute absolute difference in population by joining the reshaped 2020 blocks to the 2010 blocks 
projection <- "+proj=utm +zone=10 +ellps=GRS80 +datum=NAD83 +units=ft +no_defs" # from Ch. 2.1
sfc_pop_2010_2020 <- sfc_blocks_2010 %>% 
  left_join(sfc_blocks_2020_reshaped) %>% 
  mutate(
    absolute_pop = (pop_2020 - pop_2010) / as.numeric(st_area(.)) * 43560
  )
```

Commented out remaining code so that the project could still knit despite the errors. 
```{r}

# Attempting to use the absolute value of the change in population data from the main dataframe


max <- max(abs(sfc_pop_2010_2020$absolute_pop)) #
absolute_pal <- colorNumeric(
  palette = "PiYG",
  domain = c(-9440, 4300) # examined absolute_pop column of sfc_pop_2010_2020. 
)

# Error in Math.data.frame(sfc_pop_2010_2020 %>% select(absolute_pop)) : non-numeric-alike variable(s) in data frame: geometry

# Why is absolute_pop a geometry?? 
```

```{r}
leaflet(sfc_pop_2010_2020) %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addPolygons(
    fillColor = ~absolute_pal(absolute_pop), # what goes here?
    label = ~paste0(round(absolute_pop)),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.75,
    weight = 1
  ) %>% 
  addLegend(
    pal = absolute_pal,
    values = ~absolute_pop, # Keep scale equidistant to 0. 
    title = "Total population change in SF County from 2010 to 2020"
  )
```

This data is from the 2010 and 2020 Census, and so one assumption I made in my analysis is that all the residents living in San Francisco filled out the Census. This analysis is also assuming that the political climate surrounding the 2020 Census didn't impact data collection, which isn't necessarily a safe assumption given the partisan interference and operational issues facing the 2020 Census. 
