---
title: "DavidLudeke_A2"
author: "David Ludeke"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(tidyverse)
library(sf)
library(tigris) 
library(censusapi)
library(mapview)
library(leaflet)
```

```{r}
# Using developer version of tigris so I can grab 2020 block shapefiles correctly. 
library(devtools) 
install_github('walkerke/tigris')
library(tigris)
```

```{r}
# Requested my own Census API key from the Census. 
Sys.setenv(CENSUS_KEY="46c0518594422fc024772060dae78c2c2ffcc89d")
```

```{r}
# Grab 2020 Census data
dec_vars_2020 <-
  listCensusMetadata(
    name = "2020/dec/pl",
    type = "variables")
sfc_pop_2020 <-
  getCensus(
    name = "dec/pl",
    vintage = 2020,
    region = "block:*", 
    regionin = "state:06+county:075", # State:06 = California, County:075 = San Francisco (my hometown!)
    vars = "P1_001N") %>% # Variable code for TOTAL population
  transmute(
    GEOID20 = block,
    pop_2020 = P1_001N # Variable code for TOTAL population
  )

# Prepare 2020 census blocks for my chosen county in California.
blocks_2020 <- blocks("CA", "San Francisco", year = 2020, progress_bar = F)
```

```{r}
# Grab 2010 Census data
dec_vars_2010 <-
  listCensusMetadata(
    name = "2010/dec/pl/", 
    type = "variables"
  )
sfc_pop_2010 <-
  getCensus(
    name = "dec/pl",
    vintage = 2010,
    region = "block:*", 
    regionin = "state:06+county:075",
    vars = "P001001") %>% # Variable code for TOTAL population in the 2010 census.
  transmute(
    GEOID10 = block,
    pop_2010 = P001001
  )

# Prepare 2010 census blocks for my chosen county in California. 
blocks_2010 <- blocks("CA", "San Francisco", year = 2010, progress_bar = F)
```

Create a map of the blocks for San Francisco County based on the geometries from the 2010 Census blocks. 

```{r}
# Download the Census Designated Place shapefile for San Franciscio
sfc_boundary <- places("CA", progress_bar = F) %>% 
  filter(NAME == "San Francisco")

# Define set of SF County blocks from 2010 based on which have centroids within the boundary.
sfc_blocks_2010 <- blocks_2010 %>% 
  st_centroid() %>% 
  .[sfc_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(blocks_2010 %>% select(GEOID10)) %>% 
  st_as_sf() %>% 
  select(GEOID10) %>% 
  left_join(sfc_pop_2010) # Add in population data for each block in 2010

# Create map of boundary for San Francisco blocks 
mapview(sfc_boundary) + mapview(sfc_blocks_2010, alpha.regions = 0, color = "orange", lwd = 2)
```

```{r}
# Get all 2020 census blocks that intersect with the 2010 blocks.
sfc_blocks_2020 <- blocks_2020 %>%
  .[sfc_boundary, ] %>% 
  select(GEOID20) %>% 
  left_join(sfc_pop_2020)
```

```{r}
# Spatial subsetting method shown in Chapter 2.4.
sfc_blocks_2020_intersect <- sfc_blocks_2020 %>% 
  st_transform(26910) %>% 
  mutate(original_area = st_area(.)) %>% 
  st_intersection( # "Cookie-cutter" procedure
    sfc_blocks_2010 %>%
      select(GEOID10) %>% 
      st_transform(26910)
  ) %>% 
  mutate(
    leftover_area = st_area(.),
    perc_area = as.numeric(leftover_area / original_area),
    pop_2020 = pop_2020 * perc_area
  )
```

```{r}
# Group and summzarize to get the 2020 population counts within the GEOID10 boundaries. 
sfc_blocks_2020_reshaped <- sfc_blocks_2020_intersect %>%
  st_drop_geometry() %>% 
  group_by(GEOID10) %>% 
  summarize(
    pop_2020 = sum(pop_2020, na.rm=T) %>% round()
  )
```

Lastly, I join this to `sfc_blocks_2010` and compute the absolute difference. To compute density in people per acre, I transform to a planar coordinate system with units in ft and do the appropriate math. Note that `st_area()` produces results in a "unit" class, which can be converted to regular numbers using `as.numeric()`.

```{r}
projection <- 26910
sfc_pop_2010_2020 <- sfc_blocks_2010 %>% 
  left_join(sfc_blocks_2020_reshaped) %>% 
  mutate(
    absolute_pop = (pop_2010 - pop_2020) / as.numeric(st_area(.)) * 43560
  )
```

```{r}
# Attempting to use the absolute value of the change in population data from the main dataframe
max <- max(abs(sfc_pop_2010_2020 %>% select(absolute_pop)))
absolute_pal <- colorNumeric(
  palette = "PiYG",
  domain = c(-max,max)
)
# Error in Math.data.frame(sfc_pop_2010_2020 %>% select(absolute_pop)) : non-numeric-alike variable(s) in data frame: geometry

# Why is absolute_pop a geometry?? 
```

```{r}
leaflet() %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addPolygons(
    # fillColor = ~absolute_pal(___), # what goes here?
    # label = ~round(___),
    color = "white",
    opacity = 0.5,
    fillOpacity = 0.75,
    weight = 1
  ) %>% 
  addLegend(
    pal = absolute_pal,
    value = max, # Keep scale equidistant to 0. 
    title = "Total population change in SF County from 2010 to 2020"
  )
```


