---
title: "DavidLudeke_A5"
author: "David Ludeke"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
library(tidyverse)
library(censusapi)
library(sf)
library(tigris)
library(mapview)

library(readxl)
```

```{r}
temp <- tempfile()
download.file("https://oehha.ca.gov/media/downloads/calenviroscreen/document/calenviroscreen40resultsdatadictionaryf2021.zip",destfile = temp)

unlink(temp)

# for Mac Users, put the xlsx file in your working directory
ces4 <- read_excel("calenviroscreen40resultsdatadictionary_F_2021.xlsx")
```

```{r}
bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

ca_tracts <- tracts("CA", cb = T, progress_bar = F)
```

```{r}
ces4_clean <- ces4 %>% 
  dplyr::select(!ends_with("Pctl")) %>% 
  filter(`California County` %in% bay_county_names)

sum(is.na(ces4_clean$Poverty))

ces4_clean %>% 
  # .$Poverty %>%
  pull(Poverty) %>% 
  is.na() %>% 
  sum()

ca_tracts <- tracts("CA", progress_bar = F)

ces4_map <- ces4_clean %>% 
  filter(`California County` %in% bay_county_names) %>% 
  left_join(
    ca_tracts %>% 
      transmute(GEOID = GEOID %>% as.numeric()),
    by = c("Census Tract" = "GEOID")
  ) %>% 
  st_as_sf()

mapview(ces4_map, zcol = "Asthma")
```
Asthma data in CalEnviroScreen 4.0 is measured as the age-adjusted rate of Emergency Department (ED) visits for asthma per 10,000 people. This data is averaged over 2015-2017. As seen in the map, the highest rates of Asthma are found in the lighter green areas, including most of Vallejo, Oakland, Alameda/San Leandro, and the Bayview neighborhood in San Francisco. Particularly high-rate areas are Alameda and Vallejo. 


```{r}
# Repeat process with PM2.5
mapview(ces4_map, zcol = "PM2.5")

```

PM2.5 data in CalEnviroScreen 4.0 is measured as the annual mean concentration (which is a weighted average of measured monitor concentrations and satellite observations) in micrograms per meters cubed. The data is averaged over three years, from 2015 to 2017. The highest concentrations of PM2.5 in the Bay Area can be found in Oakland and Napa. However, most of San Francisco and the surrounding cities on the North, East, and South Bays have similar concentrations between 8 and 9 µg/m^3. 


```{r}
# Create line of best fit between Asthma and PM2.5
slope <- 0
yintercept <- mean(ces4_clean$Asthma)

best_fit_candidate <- slope * ces4_clean$PM2.5 + yintercept 
```

```{r}
# Create scatterplot
ggplot( 
  data = ces4_clean,
  aes(
    x = PM2.5,
    y = Asthma
  )
) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_line(
    aes(
      x = ces4_clean$PM2.5,
      y = best_fit_candidate 
    )
  )
```

The best fit line doesn't appear to accurately represent the data. Although there is a general positive correlation between Asthma and PM2.5, there is a huge spike in the data between 8 and 9 micrograms per meters cubed of PM2.5 that is cannot be reflected by a straight best fit line. 

```{r}
model <- lm(Asthma ~ PM2.5, ces4_clean)

summary(model)
```

An increase of 1 microgram per meter cubed of PM2.5 is associated with an increase of 19.862 in Asthma per 10000 people; 9.6% of the variation in Asthma is explained by the variation in PM2.5. This linear regression implies a correlation between Asthma and PM2.5. However, the relatively high residuals suggest the distribution might not be well-represented by the linear regression.  



## 6.5 points:
```{r}
plot(density(residuals(model)))
```
The residual distribution is reveals that there is a high density of negative residuals, since the mean of the distribution is somewhere around -20. The distribution is very right skewed, suggesting we may need a non-linear regression model. 

# Log transformation: 

```{r} 
# step 2 w/ log transformation scatter plot with logistic model
ggplot( 
  data = ces4_clean,
  aes(
    x = PM2.5,
    y = log(Asthma)
  )
) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_line(
    aes(
      x = ces4_clean$PM2.5,
      y = best_fit_candidate
    )
  )
```
A log transformation of the y-axis allows for a closer best fit line for the data.

```{r}
# step 3 with log transformation
log_model <- lm(log(Asthma) ~ PM2.5, ces4_clean)

summary(log_model)
```
An increase of 1 microgram per meter cubed of PM2.5 is associated with an increase of 1.43 in Asthma per 10000 people; 10.0% of the variation in Asthma is explained by the variation in PM2.5. This log transformation regression implies a correlation between Asthma and PM2.5. The residual standard error is 0.66, indicating a relatively close best-fit line. 


## 7 points:
```{r}
plot(density(residuals(log_model)))
```
This density plot for the log transformed data shows the residuals are much closer to 0. There is still a slight left skew, and there are two peaks on either side of 0, so while it isn't a perfect fit, it is closer than the un-transformed regression. 



# Residuals map:
```{r}
residuals <- log_model %>% 
  .$residuals %>% 
  as.data.frame()

ces4_res <-
  ces4_clean %>% filter(
    !is.na(PM2.5),
    !is.na(Asthma)) %>% 
  mutate(residuals = residuals(log_model))
```

```{r}
ces4_res_map <- ces4_res %>% 
  filter(`California County` %in% bay_county_names) %>% 
  select(`Census Tract`, residuals) %>% 
  left_join(
    ca_tracts %>% 
      transmute(GEOID = GEOID %>% as.numeric()),
    by = c("Census Tract" = "GEOID")
  ) %>% 
  st_as_sf()
```

```{r}
mapview(ces4_res_map, zcol = "residuals")
```
The two census tracts with the most negative residual are the ones for Stanford! In the context of Asthma estimation, a negative residual means that there is a over-estimation of Asthma rates based on the PM2.5 levels. This could be the case for this particular census tract because Stanford's population is composed of students from all of California, the country, and the world, and so they grew up in areas with different PM2.5 levels and thus are likely to have different Asthma rates.  


