---
title: "DavidLudeke_A5"
author: "David Ludeke"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```


```{r}
library(tidyverse)
library(censusapi)
library(sf)
library(tigris)
library(mapview)

library(readxl)
```

```{r}

temp <- tempfile()
download.file("https://oehha.ca.gov/media/downloads/calenviroscreen/document/calenviroscreen40resultsdatadictionaryf2021.zip",destfile = temp)

unlink(temp)

# for Mac Users, put the xlsx file in your working directory
ces4 <- read_excel("calenviroscreen40resultsdatadictionary_F_2021.xlsx")
```

```{r}
bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

ca_tracts <- tracts("CA", cb = T, progress_bar = F)
```

```{r}
ces4_clean <- ces4 %>% 
  dplyr::select(!ends_with("Pctl")) %>% 
  filter(`California County` %in% bay_county_names)

sum(is.na(ces4_clean$Poverty))

ces4_clean %>% 
  # .$Poverty %>%
  pull(Poverty) %>% 
  is.na() %>% 
  sum()

ca_tracts <- tracts("CA")

ces4_map <- ces4_clean %>% 
  filter(`California County` %in% bay_county_names) %>% 
  left_join(
    ca_tracts %>% 
      transmute(GEOID = GEOID %>% as.numeric()),
    by = c("Census Tract" = "GEOID")
  ) %>% 
  st_as_sf()

mapview(ces4_map, zcol = "Asthma")
```

```{r}
# Repeat process with PM2.5
mapview(ces4_map, zcol = "PM2.5")

```

```{r}
# Create line of best fit between Asthma and PM2.5
slope <- 0
yintercept <- mean(ces4_clean$Asthma)

best_fit_candidate <- slope * ces4_clean$PM2.5 + yintercept 
```

```{r}
# Create scatterplot
ggplot( 
  data = ces4_clean,
  aes(
    x = PM2.5,
    y = Asthma
  )
) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_line(
    aes(
      x = ces4_clean$PM2.5,
      y = best_fit_candidate 
    )
  )
```

The best fit line doesn't appear to accurately represent the data. Although there is a general positive correlation between Asthma and PM2.5, there is a huge spike in the data between 8 and 9 micrograms per meters cubed of PM2.5 that is cannot be reflected by a straight best fit line. 

```{r}
model <- lm(Asthma ~ PM2.5, ces4_clean)

summary(model)
```

An increase of 1 microgram per meter cubed of PM2.5 is associated with an increase of 19.862 in Asthma per 10000 people; 9.6% of the variation in Asthma is explained by the variation in PM2.5.
